---
title: "Final Project"
output:
  pdf_document: default
  html_document: default
date: "2023-11-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#load data
```{r}
library(lme4)
library(lmerTest)
```

#df
```{r}
Tester.name <- c("Joe", "Alex","Laura", "Jade")
    Testers <- as.factor(rep(Tester.name,1, each =9))
Water.Level <- as.factor(rep(2:4, 4, each = 3)) 
    #W1 = 2 , W2 = 3 , W3 = 4
Additive <- c("Honey", "Lemon", "Sugar")
    Additives <- as.factor(rep(Additive,12))

     #A1 = honey , A2 = lemon , A3 = sugar
Ratings <- c(4,7,6,2,7,6,2,8,7,7,4,5,8,2,5,7,4,5,7,6,8,8,2,7,9,7,7,4,1,9,8,2,9,8,1,6)
Block <- as.factor(rep(1:2,1,each=18))
    #Block 1 = Olivia O. , Block 2 = Olivia R.
    
Apple.Juice.Data <- data.frame(Block, Water.Level, Additives, Testers, Ratings)
```

```{r}
set.seed(2023)
```



#Water.Level * Additives + (1 | Block)
```{r}
res <- lmer(Ratings ~ Water.Level*Additives+(1 | Block),data=Apple.Juice.Data)
anova(res)
summary(res)
```

##assumption graphs
```{r}
plot(res)
qqnorm(residuals(res))
qqline(residuals(res))
```
#Additives + Water.Level * Additives + (1 | Block)
```{r}
res2 <- lmer(Ratings ~ Additives + Water.Level*Additives+(1 | Block),data=Apple.Juice.Data)
anova(res2)
summary(res2)
```


##assumption graphs
```{r}
plot(res2)
qqnorm(residuals(res2))
qqline(residuals(res2))
```
#Additives + Water.Level + Water.Level * Additives + (1 | Block)
```{r}
res3 <- lmer(Ratings ~ Additives + Water.Level+(1 | Block),data=Apple.Juice.Data)
anova(res3)
summary(res3)
```
##assumption graphs
```{r}
plot(res3)
qqnorm(residuals(res3))
qqline(residuals(res3))
```

#Block+Additives + Water.Level + Water.Level * Additives +(1 | Block:Water.Level)
```{r}
res4 <- lmer(Ratings ~ Block+Additives+Water.Level +(1 | Block),data=Apple.Juice.Data)
anova(res4)
summary(res4)
```
##assumption graphs
```{r}
plot(res4)
qqnorm(residuals(res4))
qqline(residuals(res4))
```


#Block + Additives + Water.Level + (1 | Block)
```{r}
res6 <- lmer(Ratings ~ Block+ Additives+Water.Level +(1 | Block),data=Apple.Juice.Data)
anova(res6)
summary(res6)
```
##assumption graphs
```{r}
plot(res6)
qqnorm(residuals(res6))
qqline(residuals(res6))
```

```{r}
AIC(res, res2, res3, res4, res5, res6)
```

```{r}
anova(res2,res4)
```



#Interaction Plot
```{r}
interaction.plot(x.factor = Additives, 
                            trace.factor = Water.Level, 
                            response = Ratings, 
                            main = "Interaction Plot for Ratings",
                            type = "o",
                            pch=c(8,17,19),
                            lwd=2,
                            col = c("darkgoldenrod1", "cornflowerblue","darkgreen"),
                            ylim=c(2,7),
                            legend = FALSE)
legend("bottomright",
       c("2","3","4"),
       bty="n",
       lty=c(1,2),
       lwd=2,
       col=c("darkgoldenrod1", "cornflowerblue","darkgreen"), 
       pch=c(8,17,19),
       inset = .02,
       title="Cups of Water Boiled")
                  
```

```{r}
par(mfrow = c(2, 2))
plot(res4)
```

```{r}
plot(res4)
```

```{r}
anova(res3,res2)
```

$H_0:\alpha_i=0$ for all i

$H_A:\alpha_i/neq0$ for at least one i

$H_0: \beta_j=0$ for all j

$H_A: \beta_j\neq0$ for at least one j



```{r}

library(devtools)
library(redres)
# creates a plot of the conditional studentized residuals versus the fitted values
plot_redres(res3, type = "std_cond")
# creates a residual quantile plot for the error term
plot_resqq(res3)
# creates normal quantile plots for each random effect
plot_ranef(res3)

```

